---
title: "Pacific giant salamanders (Dicamptodon tenebrosus) in Mack Creek"
author: "Anne-Marie Parkinson and Hannah Garcia-something with a W"
date: "November 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load packages ------------------------------------------------------------------

library (tidyverse)
library (janitor)
library (dplyr)
library (kableExtra)

# add data -------------------------------------------------------------------------

mack_creek_verts <- read_csv("mack_creek_vertebrates.csv") %>% 
  clean_names()

```


### Introduction

url for picture: http://www.californiaherps.com/salamanders/pages/d.tenebrosus.html
photo credit: Gary Nafis
### Data and Methods

### Results A: Annual salamander counts in old growth and clear cut sections of Mack Creek

```{r}
# subdata: remove unneccessary columns and keep only salamander observations
salamander_counts_data <- mack_creek_verts %>% 
  filter (species == "DITE") %>% 
  select (year, section, species)

# count number of salamanders in each section (CC vs OG) for every year
salamander_counts <- salamander_counts_data %>% 
  group_by(year, section) %>% 
  count(species)

# plot - not the right graph type
ggplot(salamander_counts, aes (x=year, y=n)) +
  geom_col(aes(fill=section))


```


### Results B: Table of 2017 salamander counts by channel classification (pool, cascades and side-channel) in old growth and clear cut sections of Mack Creek. 

```{r}
#subdata: only include 2017 and salamader observations, exclude isolated pools from data
salamander_channel_2017 <- mack_creek_verts %>% 
  filter (year == "2017") %>% 
  filter (species == "DITE") %>% 
  filter (unittype != "IP") %>% 
  mutate(channel_type= case_when(
    unittype %in% "C" ~ "Cascade",
    unittype %in% "P" ~"Pool",
    unittype %in% "SC" ~ "Side Channel"
  )) %>% 
  select (species, channel_type, section)

# counts
salamander_channel_counts_2017 <- salamander_channel_2017 %>% 
  group_by(channel_type, section) %>% 
  count(species) %>% 
  select (channel_type, section, n)

# create contigency table 
salamander_table_2017 <- salamander_channel_counts_2017 %>% 
  pivot_wider(names_from = section, values_from = n) 
  

# add proportions to contigency table
salamander_table_counts_proportions_2017 <- salamander_table_2017 %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting(digits=1) %>% 
  adorn_ns(position = "front")

#  markdown ready table
salamander_table_counts_proportions_2017 %>% 
  kable (col.names = c("Channel Type", "Clear Cut Forest", "Old-Growth Forest")) %>% 
  kable_styling()

  
```
not sure why getting that error message. its not affecting the table counts/proportions (ex/when change position from front to rear dont get the error message)

### Results c: 

### Results D: 

### Results E:

### Summary

### References 
