---
title: "Pacific giant salamanders in Mack Creek"
author: "Anne-Marie Parkinson and Hannah Garcia-Wickstrum"
date: "November 24, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE)
```

```{r}
# load packages ------------------------------------------------------------------

library (tidyverse)
library (janitor)
library (dplyr)
library (kableExtra)

# add data -------------------------------------------------------------------------

mack_creek_verts <- read_csv("mack_creek_vertebrates.csv") %>% 
  clean_names()

```


### Introduction

url for picture: http://www.californiaherps.com/salamanders/pages/d.tenebrosus.html
photo credit: Gary Nafis
### Data and Methods

### Results

#### Results A: Annual salamander counts in old growth and clear cut sections of Mack Creek

```{r}
# subdata: remove unneccessary columns and keep only salamander observations
salamander_counts_data <- mack_creek_verts %>% 
  filter (species == "DITE") %>% 
  select (year, section, species)

# count number of salamanders in each section (CC vs OG) for every year
salamander_counts <- salamander_counts_data %>% 
  group_by(year, section) %>% 
  count(species)

# plot - not the right graph type
# ggplot(salamander_counts, aes (x=year, y=n)) +
#   geom_col(aes(fill=section)) 

ggplot(salamander_counts, aes(fill = section, x = year, y = n)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_fill_manual("legend", values = c("CC" = "orange", "OG" = "seagreen")) +
  scale_x_continuous(lim = c(1993, 2017),
                     expand = c(0, 0),
                     breaks = seq(1993, 2017, by = 2)) +
  labs(x = "year", y = "salamander counts") +
  ggtitle("Salamander counts in old growth and clear cut sections")
  
# what do you think about the grouped bar? and the colors? do you think we should have the legend list out the full name instead of the acronym? or we can just clearly state in title and figure caption?
```
***Figure #.** Salamander counts in clear cut (orange) and old growth (green) sections of Mack Creek. From 1993 to 2013, there were almost always more salamanders counted in old growth sections. From 2013 to 2017, salamander counts were higher in clear cut sections. Data: Stanley Gregory, Andrews Forest LTER Sites.* **Needs editing and better description.**

### Results B: Table of 2017 salamander counts by channel classification (pool, cascades and side-channel) in old growth and clear cut sections of Mack Creek. 

```{r}
#subdata: only include 2017 and salamader observations, exclude isolated pools from data
salamander_channel_2017 <- mack_creek_verts %>% 
  filter (year == "2017") %>% 
  filter (species == "DITE") %>% 
  filter (unittype != "IP") %>% 
  mutate(channel_type= case_when(
    unittype %in% "C" ~ "Cascade",
    unittype %in% "P" ~"Pool",
    unittype %in% "SC" ~ "Side Channel"
  )) %>% 
  select (species, channel_type, section)

# counts
salamander_channel_counts_2017 <- salamander_channel_2017 %>% 
  group_by(channel_type, section) %>% 
  count(species) %>% 
  select (channel_type, section, n)

# create contigency table 
salamander_table_2017 <- salamander_channel_counts_2017 %>% 
  pivot_wider(names_from = section, values_from = n) 
  

# add proportions to contigency table
salamander_table_counts_proportions_2017 <- salamander_table_2017 %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting(digits=1) %>% 
  adorn_ns(position = "front")

#  markdown ready table
salamander_table_counts_proportions_2017 %>% 
  kable (col.names = c("Channel Type", "Clear Cut Forest", "Old-Growth Forest")) %>% 
  kable_styling()

  
```
not sure why getting that error message. its not affecting the table counts/proportions (ex/when change position from front to rear dont get the error message)

### Results C: 

### Results D: 

### Results E:

### Summary

### References 

**Gregory S. V. 2016**. Aquatic Vertebrate Population Study in Mack Creek, Andrews Experimental Forest, 1987 to present. Environmental Data Initiative. https://doi.org/10.6073/pasta/5de64af9c11579266ef20da2ff32f702. Dataset accessed 11/25/2019.
